Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-01-12 grove-paper.tex
Add related work for CRDTs

@@

 commutative replicated data types CmRDT
 convergent replicated data types, or CvRDTs
 Delta state CRDTs[12][13] (or simply Delta CRDTs

@@

PN-Set

@@

textcomp Genealogical Symbols
\textborn
\textdied

>>>>

Lots of list-of-chars or list-of-list-of-chars (we ignore these except to discuss them here)

@@

\url{https://github.com/PsychoLlama/graph-crdt}
 - Graph CRDT
 - Uses a LWW-E-Set

\url{https://martin.kleppmann.com/2020/07/06/crdt-hard-parts-hydra.html} (overview talk)
 We don't have interleaving problems because
   - everything is relative to a specific ID not a position
   - we don't try to auto resolve
   - we are tree not list
   Part three: moving sub tree
   - Last parent writter wins (prevents cycles)
     Equivalent to us if we filter multiparent edges, different for cycles, no way to delete?

...
 commutative replicated data types CmRDT
 convergent replicated data types, or CvRDTs
 Delta state CRDTs[12][13] (or simply Delta CRDTs

The Causal Graph CRDT for Complex Document Structure

@@

G-Set
2P-Set
PN-Set
LWW-element-Register
OR-Set
MV-Register: Multi-Value Register
U-Set (this is what we are for edges, not OR-set due to duplicates)
Add-Remove Partial Order data type
  2P-Set for vertices, and a G-Set for edges.

A comprehensive study of Convergent and Commutative Replicated Data Types (2011)
  \url{https://hal.inria.fr/inria-00555588/document}

CRDTs: Consistency without concurrency control
  \url{https://arxiv.org/abs/0907.0929}

@@

https://crdt.tech/papers.html
[35]
    Mahsa Najafzadeh, Marc Shapiro, and Patrick Eugster. Co-design and verification of an available file system. In 19th International Conference on Verification, Model Checking, and Abstract Interpretation, VMCAI 2018, pages 358--381. Springer LNCS volume 10747, January 2018. [ bib | DOI | .pdf ]

        Keywords: trees

[41]
    Martin Kleppmann and Alastair R Beresford. A conflict-free replicated JSON datatype. IEEE Transactions on Parallel and Distributed Systems, 28(10):2733--2746, April 2017. [ bib | DOI | arXiv ]

        Keywords: trees

[67]
    Vinh Tao, Marc Shapiro, and Vianney Rancurel. Merging semantics for conflict updates in geo-distributed file systems. In 8th ACM International Systems and Storage Conference, SYSTOR 2015. ACM, May 2015. [ bib | DOI | .pdf ]

        Keywords: trees

[95]
    Mehdi Ahmed-Nacer, Stéphane Martin, and Pascal Urso. File system on CRDT. Research Report RR-8027, INRIA, July 2012. [ bib | arXiv | http ]

        Keywords: trees

[103]
    Stéphane Martin, Pascal Urso, and Stéphane Weiss. Scalable XML collaborative editing with undo. In On the Move to Meaningful Internet Systems (OTM), pages 507--514. Springer LNCS volume 6426, October 2010. [ bib | DOI | arXiv ]

        Keywords: trees, undo
